Backpropagating Gravitational wave events

### 2022-04-13 src/ms.tex
Move root finding around

\section{Result}
\label{sec:result}

@@

By leveraging gradient-based optimization strategy, obtaining posterior samples in the evolutionary parameter space is much more efficient than rejection sampling.
In each optimization step during the root-finding process, the current state of the guess and its gradient provide useful information about where the root might be.
In contrast to sampling algorithm, which discard the sample whenever it is not accepted.
This greatly benefit the efficiency of the root-finding process.
On the other hand, this method is not a sampling algorithm, but a coordinate transform of the posterior samples.
One should pay extra caution in interpreting the posterior samples in the evolutionary parameter space.
Since posterior samples in the observable space can in principle have multiple roots in the evolutionary parameter space,
the relatively weighting of these roots becomes ambiguous when the number of roots is large.
For example, should we consider roots that are close to each other a single root or multiple roots?
If we have two clusters of roots, the relative number of roots within each cluster would determine the relative weight of each mode.
Note that since they both map to the same observable posterior sample within the acceptance threshold, all weighting assigned are equally valid from a "matching data" standpoint.
This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root,
and we have checked the posterior samples produced by this method for events discussed in this paper agree with direct sampling using a uniform prior.
This is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
We leave a detail comparison between this method and direct sampling to future studies.

>>>>

On top of using the raw result from the root-finding process, we also perform MCMC sampling \footnote{We use sampling package \texttt{emcee}\kw{cite} for the MCMC.} using the roots to initialize the MCMC chains.
The two results agree well with each other.
This should eliminate all concern related to the ambiguity in interpreting the roots because MCMC result should not depend on the initialization.less
\section{Result}
\label{sec:result}

@@

By leveraging gradient-based optimization strategy, obtaining posterior samples in the evolutionary parameter space is much more efficient than rejection sampling.
In each optimization step during the root-finding process, the current state of the guess and its gradient provide useful information about where the root might be.
In contrast to sampling algorithm, which discard the sample whenever it is not accepted.
This greatly benefit the efficiency of the root-finding process.
On the other hand, this method is not a sampling algorithm, but a coordinate transform of the posterior samples.
One should pay extra caution in interpreting the posterior samples in the evolutionary parameter space.
Since posterior samples in the observable space can in principle have multiple roots in the evolutionary parameter space,
the relatively weighting of these roots becomes ambiguous when the number of roots is large.
For example, should we consider roots that are close to each other a single root or multiple roots?
If we have two clusters of roots, the relative number of roots within each cluster would determine the relative weight of each mode.
Note that since they both map to the same observable posterior sample within the acceptance threshold, all weighting assigned are equally valid from a "matching data" standpoint.
This leave the interpretation of the posterior in the evolutionary parameter space ambiguous.
In our case, we found empirically each posterior sample may at most only correspond to a handful of roots (most of the posterior samples have a unique root.), therefore we choose to assign equal weight to each root.

The potential reason for the good agreement between MCMC and the root finding is due to the behavior of posterior in the evolutionary parameter space are larger determined by the posterior distribution in the observable space instead of multiple possible roots per posterior sample.
We have also compared two MCMC results, one is initialized using the roots and another is initialized using samples drawn from a uniform prior.
We found empirically the MCMC result using the roots as initialization converges at least an order of magnitude faster than the uniform prior initialization.
This makes sense since the roots basically eliminate the long burn-in needed to go from random location in the evolutionary parameters to the most probable region.
We leave a detail comparison between this method and direct sampling to future studies.
